Tight Bounds for Online Balanced Graph Partitioning

### 2020-05-08 paper/disc20clustering.tex

In this section we consider a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
To avoid infinite cost, each communicating pair must be collocated.
We study instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters.
The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.

For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
The lower bound can be easily translated to the partitioning model (studied in Section~\ref{sec:part}) with finite request sequence.
At the end of this section, we discuss an asymptotically optimal upper bound.

@@

	As \TAlg is component respecting, it never increases any counter above $\alpha$.
	We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
	Let $\sigma$ be the input sequence that arrived during the phase.
	Let $\sigma_{cost}$ be the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incurred a cost for \TAlg).
	In our analysis, we partition $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
	The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
	The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.

@@

	Now we lower-bound the cost of $\OPT$.
	By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on these input sequences (defined with respect to components of \TAlg in this phase).
	By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$ in this phase.

>>>>

In this section we consider the learning variant of Online Balanced Partitioning.
For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
The lower bound can be easily translated to the partitioning model (studied in Section~\ref{sec:part}) with finite request sequence.
At the end of this section, we discuss an asymptotically optimal upper bound.

@@

	As \TAlg is component respecting, it never increases any counter above $\alpha$.
	We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
	By $\sigma$ we denote the input sequence that arrived during the phase.
	By $\sigma_{cost}$ we denote the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incur cost for \TAlg).
	In our analysis, we partition $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
	The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
	The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.

@@

	Now we lower-bound the cost of $\OPT$.
	By $\OPT(\sigma_I)$ and $\OPT(\sigma_E)$ we denote the cost of $\OPT$ on requests from sequences $\sigma_I$ and $\sigma_E$, respectively.
	Note that these costs are defined with respect to components of \TAlg in this phase.
	By $\OPT(M)$ we denote the cost of migrations performed by $\OPT$ in this phase.
